---
title: "stochastic-population-model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stochastic-population-model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{tibble}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(villager)
```

## Background
This is an example of how to build a village with a stochastic population model. At each year, the population changes to a random value.

## Creating the Model
The first step is to define the function that will change the population at each year. Recall that it _must_ take `currentState`, `previousState`, and `modelData` as parameters.

``` {R}
newPopulation <- function(currentState, previousState, modelData, population_manager) {
  # Set the stock to a random number
  currentState$fishStock <- runif(1,10,15)
}

```

## Creating the Village
The village creation is done the standard way. An initial condition is created using the default state object and is added to the village with the model.
``` {R}
# Create the village
library(tibble)
village_initial_condition = VillageState$new(fishStock=10)
village = BaseVillage$new(name="Random fish stock Village",
                          initialState=village_initial_condition,
                          models = newPopulation)

mayanSimulation <- Simulation$new(length=10, #days
                                   villages=c(village))

# Call run_model to start the simulation
mayanSimulation$run_model()
mayanSimulation$show_results(dependent_variable="fishStock")
```
