---
title: "village"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{village}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(villager)
```

## The `village` Class

The `village` class represents a village. Internally, the village has 


### Initializing the Village
The `village` object relies on an initial condition, which is a VillageState class.

```{R}
# Create the village with the initial condition
village = BaseVillage$new()
```


```{R}
new_population <- function(currentState, previousState, modelData) {
  # Set the population to 10
  currentState$population <- 10
}

# Create the village with the initial condition
village = BaseVillage$new(models=new_population)
```



### Propagating Through Time
The village `propagate` method advances the village a single time step. In doing so, it calls the user-defined models. It also creates the new
state for the new time. This state and the previous are passed to the user's models where they can be changed according to the model's rules


### Adding Trade Partners
Villages can be connected to other villages for trade by calling `add_trade_partner(newTradePartner, addBack=TRUE)`. Internally, this appends a reference to `newTradePartner` to the list of villages that can be traded with: `tradePartners`.

Since trade is generally bi-directional between villages but not always (it should be possible for 'donations'), `addBack` can be used to make trading one directional.






